Minimax-optimal Laplacian Eigenmaps regression over Sobolev Spaces with Neighborhood Graphs

### 2021-04-27 manuscripts/laplacian_eigenmaps/appendix.tex
Massive update to Laplacian Eigenmaps draft. Too many changes to list, but all generally responding to Ryan's comments.

\begin{lemma}
	\label{lem:fixed_graph_estimation}
	For any $s \in \mathbb{N} \setminus \{0\}$, the Laplacian eigenmaps estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
	\begin{equation*}
	\|\wh{f} - f_0\|_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{5K}{n},
	\end{equation*}
	with probability at least $1 - \exp(-K)$. 
\end{lemma}
\begin{proof}
	By the triangle inequality,
	\begin{equation}
	\label{pf:fixed_graph_estimation_1}
	\|\wh{f} - f_0\|_n^2 \leq 2\Bigl(\|\mathbb{E}\wh{f} - f_0\|_n^2 + \|\wh{f} - \mathbb{E}\wh{f}\|_n^2\Bigr).
	\end{equation}
	The first term in~\eqref{pf:fixed_graph_estimation_1} (bias) is non-random, since the design is fixed. The expectation $\mathbb{E}\wh{f} = \sum_{k = 1}^{K} \dotp{v_k}{f_0}_n v_k$, so that
	\begin{equation*}
	\|\mathbb{E}\wh{f} - f_0\|_n^2 = \Bigl\|\sum_{k = K + 1}^{n} \dotp{v_k}{f_0}_n v_k\Bigr\|_n^2 = \sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2.
	\end{equation*}
	In the above, the last equality relies on the fact that $v_k$ are orthonormal. Using the fact that the eigenvalues are in increasing order, we obtain that
	\begin{equation*}
	\sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2 \leq \frac{1}{\lambda_K^s} \sum_{k = K + 1}^n \lambda_k^s \dotp{v_k}{f_0}_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s}.
	\end{equation*}

	The second term in~\eqref{pf:fixed_graph_estimation_1} (variance) is random. Again using the orthonormality of the eigenvectors $v_k$, we have
	\begin{equation*}
	\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{\varepsilon}_n^2.
	\end{equation*}
	Thus $\|\wh{f} - \mathbb{E}\wh{f}\|_n^2$ is equal to $1/n$ times a $\chi^2$ distribution with $K$ degrees of freedom. Consequently, it follows from a result of \citep{laurent00} that
	\begin{equation*}
	\Pbb\Bigl(\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 \geq \frac{K}{n} + 2\frac{\sqrt{K}}{n}\sqrt{t} + \frac{2t}{n}\Bigr) \leq \exp(-t).
	\end{equation*}
	Setting $t = K$ completes the proof of the Lemma.
\end{proof}

@@

\begin{lemma}
	Suppose we observe $(Y_1,x_1),\ldots,(Y_n,x_n)$ according to~\eqref{eqn:fixed_graph_regression_model}.
	\begin{itemize}
		\item If $f_0 = 0$, then $\Ebb_0[\varphi] \leq a$.
		\item Suppose $f_0 \neq 0$ satisfies
		\begin{equation*}
		\|f_0\|_n^2 \geq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + 16\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr],
		\end{equation*}
		for some $s \in \mathbb{N}\setminus \{0\}$. Then $\Ebb_{f_0}[1 - \phi] \leq b$.
	\end{itemize}
\end{lemma}

@@

In Lemma~\textcolor{red}{(?)}, we show that the graph Sobolev semi-norm (V-statistic) decomposes into an unbiased estimate of the non-local Sobolev semi-norm (U-statistic), and a pure bias term. We also establish that the pure bias term will be small (in expectation) relative to the U-statistic whenever $\varepsilon$ is sufficiently small.
\begin{lemma}
	\label{lem:graph_seminorm_bias}
	For any $f \in L^2(\mc{X})$, the graph Sobolev semi-norm satisfies
	\begin{equation}
	\label{eqn:graph_seminorm_bias_1}
	\dotp{L_{n,\varepsilon}^sf}{f}_{n} = U_{n,\varepsilon}^{(s)}(f) + B_{n,\varepsilon}^{(s)}(f),
	\end{equation}
	such that $\mathbb{E}[U_{n,\varepsilon}^{(s)}(f)] = (n - s - 1)!/n! \cdot \dotp{L_{P,\varepsilon}^sf}{f}_P$. If additionally $f \in H^1(\mc{X};M)$ and $\varepsilon \geq n^{-1/d}$, then the bias term $B_{n,\varepsilon}^{(s)}(f)$ satisfies
	\begin{equation}
	\label{eqn:graph_seminorm_bias_2}
	\mathbb{E}\bigl[|B_{n,\varepsilon}^{(s)}(f)|\bigr] \leq \frac{C\varepsilon^2}{\delta n\varepsilon^{2 + d}}M^2.
	\end{equation}
\end{lemma}
Then~\ref{pf:graph_seminorm_ho_1} follows immediately from Lemma~\ref{lem:graph_seminorm_bias}, by Markov's inequality.
\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias}).}
We begin by introducing some notation. We will use bold notation $\bj = (j_1,\ldots,j_s)$ for a vector of indices where $j_i \in [n]$ for each $i$. We write $[n]^s$ for the collection of all such vectors, and $(n)^s$ for the subset of such vectors with no repeated indices. Finally, we write $D_if$ for a kernelized difference operator,
\begin{equation*}
D_if(x) := \bigl(f(x) - f(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
\end{equation*}
and we let $D_{\bj}f(x) := \bigl(D_{j_1}\circ \cdots \circ D_{j_s}f\bigr)(x)$.

@@

It remains to show~\eqref{eqn:graph_seminorm_bias_2}. By adding and subtracting $f(X_{\bj_1})$, we obtain by symmetry that
\begin{equation*}
\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i) = \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot \bigl(f(X_i) - f(X_{\bj_1})\bigr),
\end{equation*}
and consequently
\begin{equation*}
\Ebb\Bigl[\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i)\Bigr] \leq \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr].
\end{equation*}
In Lemma~\textcolor{red}{(?)}, we show that if $f \in H^1(\mc{X};M)$, then for any $i\bj \in [n]^{s + 1}$ which contains a total of $k + 1$ distinct indices, 
\begin{equation*}
\Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr] \leq C_1 \varepsilon^{2 + kd} M^2.
\end{equation*}
This shows us that the expectation of $|B_{n,\varepsilon}^s(f)|$ can bounded from above by the sum over several different terms, as follows:
\begin{align*}
\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] & \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2 \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \frac{1}{(n\varepsilon^d)^s}  \varepsilon^{(|i\bj| - 1)d} \\
& \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2  \sum_{k = 1}^{s - 1} \frac{(n\varepsilon^d)^k}{(n\varepsilon^d)^s}n.
\end{align*}
Finally, we note that by assumption $n\varepsilon^d \geq 1$, so that in the above sum the factor of $(n\varepsilon^d)^k$ is largest when $k = s- 1$. We conclude that
\begin{equation*}
\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] \leq C_1 (s - 1) \frac{\varepsilon^2}{n\varepsilon^{2s + d}}M^2,
\end{equation*}
which is the desired result.

@@

Now, we use the results of the preceding two sections to prove~\eqref{pf:graph_seminorm_ho_2}. We will divide our analysis in two cases, depending on whether $s$ is odd or even, but before we do this we state some facts that will be applicable to both cases. First, we  recall that $L_{P,\varepsilon}$ is self-adjoint in $L^2(P)$, meaning $\dotp{L_{P,\varepsilon}f}{g}_{P} = \dotp{f}{L_{P,\varepsilon}g}_{P}$ for all $f, g \in L^2(P)$. We will also use the fact that when $s = 1$, the non-local seminorm is a \emph{non-local Dirichlet energy},

@@

\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}^{k + 1}f}{L_{P,\varepsilon}^{k}f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}).

@@

\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_bias}}
\label{subsec:pf_kernel_smoothing_bias}

>>>>

\begin{lemma}
	\label{lem:fixed_graph_estimation}
	For any $s \in \mathbb{N} \setminus \{0\}$, the Laplacian eigenmaps estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
	\begin{equation}
	\label{eqn:fixed_graph_estimation}
	\|\wh{f} - f_0\|_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{5K}{n},
	\end{equation}
	with probability at least $1 - \exp(-K)$. 
\end{lemma}
\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_estimation}).}
	By the triangle inequality,
	\begin{equation}
	\label{pf:fixed_graph_estimation_1}
	\|\wh{f} - f_0\|_n^2 \leq 2\Bigl(\|\mathbb{E}\wh{f} - f_0\|_n^2 + \|\wh{f} - \mathbb{E}\wh{f}\|_n^2\Bigr).
	\end{equation}
	The first term in~\eqref{pf:fixed_graph_estimation_1} (approximation error) is non-random, since the design is fixed. The expectation $\mathbb{E}\wh{f} = \sum_{k = 1}^{K} \dotp{v_k}{f_0}_n v_k$, so that
	\begin{equation*}
	\|\mathbb{E}\wh{f} - f_0\|_n^2 = \Bigl\|\sum_{k = K + 1}^{n} \dotp{v_k}{f_0}_n v_k\Bigr\|_n^2 = \sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2.
	\end{equation*}
	In the above, the last equality relies on the fact that $v_k$ are orthonormal. Using the fact that the eigenvalues are in increasing order, we obtain that
	\begin{equation*}
	\sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2 \leq \frac{1}{\lambda_K^s} \sum_{k = K + 1}^n \lambda_k^s \dotp{v_k}{f_0}_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s}.
	\end{equation*}

	The second term in~\eqref{pf:fixed_graph_estimation_1} (estimation error) is random. Observe that $\dotp{v_k}{\varepsilon}_n \overset{d}{=} Z_k/\sqrt{n}$, where $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$. Again using the orthonormality of the eigenvectors $v_k$, we have
	\begin{equation*}
	\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{\varepsilon}_n^2 \overset{d}{=} \frac{1}{n}\sum_{k = 1}^{K} Z_k^2.
	\end{equation*}
	Thus $\|\wh{f} - \mathbb{E}\wh{f}\|_n^2$ is equal to $1/n$ times a $\chi^2$ distribution with $K$ degrees of freedom. Consequently, it follows from a result of \citep{laurent00} that
	\begin{equation*}
	\Pbb\biggl(\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 \geq \frac{K}{n} + 2\frac{\sqrt{K}}{n}\sqrt{t} + \frac{2t}{n}\biggr) \leq \exp(-t).
	\end{equation*}
	Setting $t = K$ completes the proof of the Lemma.

@@

\begin{lemma}
	\label{lem:fixed_graph_testing}
	Suppose we observe $(Y_1,x_1),\ldots,(Y_n,x_n)$ according to~\eqref{eqn:fixed_graph_regression_model}.
	\begin{itemize}
		\item If $f_0 = 0$, then $\Ebb_0[\varphi] \leq a$.
		\item Suppose $f_0 \neq 0$ satisfies
		\begin{equation}
		\label{eqn:fixed_graph_testing_critical_radius}
		\|f_0\|_n^2 \geq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr],
		\end{equation}
		for some $s \in \mathbb{N}\setminus \{0\}$. Then $\Ebb_{f_0}[1 - \phi] \leq b$.
	\end{itemize}
\end{lemma}
\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_testing}).}
We first compute the expectation and variance of $\wh{T}$, then apply Chebyshev's inequality to upper bound the Type I and Type II error.

\underline{\emph{Expectation}.}
Recall that $\wh{T} = \sum_{k = 1}^{K} \dotp{Y}{v_k}_n^2$. Expanding the square gives
\begin{equation*}
\Ebb[\wh{T}] = \sum_{k = 1}^{K} \Ebb[\dotp{Y}{v_k}_n^2] = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 + \Ebb[2\dotp{f_0}{v_k}_n\dotp{\varepsilon}{v_k}_n + \dotp{\varepsilon}{v_k}_n^2] = \frac{K}{n} + \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2.
\end{equation*}
Thus $\Ebb[\wh{T}] - t_a = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 - \sqrt{2K}/n \cdot \sqrt{1/a}$. Furthermore, it is a consequence of~\eqref{eqn:fixed_graph_testing_critical_radius} that 
\begin{equation}
\label{pf:fixed_graph_testing_1}
\sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 - \frac{\sqrt{2K}}{n}\sqrt{1/a} \geq \|f_0\|_n^2 - \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} - \frac{\sqrt{2K}}{n}\sqrt{1/a} \geq \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr].
\end{equation}

\underline{\emph{Variance}.}
Recall from the proof of Lemma~\ref{lem:fixed_graph_estimation} that $\dotp{\varepsilon}{v_k}_n \overset{d}{=} Z_k/\sqrt{n}$ for $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$. Expanding the square, and recalling that $\Cov[Z,Z^2] = 0$ for Gaussian random variables, we have that
\begin{equation*}
\Var\bigl[\dotp{{\bf Y}}{v_k}_n^2\bigr] = \Var\biggl[\frac{2}{n}\dotp{f_0}{v_k}_nZ_k + \frac{2}{n^2}Z_k^2\biggr] = \frac{4\dotp{f_0}{v_k}_n^2}{n} + \frac{2}{n^2}.
\end{equation*}
Moreover, since $\Cov[Z_k^2,Z_{\ell}^2] = 0$ for each $k = 1,\ldots,K$, we see that
\begin{equation*}
\Var\bigl[\wh{T}\bigr] = \sum_{k = 1}^{K} \Var\bigl[\dotp{{\bf Y}}{v_k}_n^2\bigr] = \frac{2K}{n^2} + \sum_{k = 1}^{K}\frac{4\dotp{f_0}{v_k}_n^2}{n}.
\end{equation*}

\underline{\emph{Bounds on Type I and Type II error}.}
The upper bound on Type I error follows immediately from Chebyshev's inequality.

The upper bound on Type II error also follows from Chebyshev's inequality. We observe that~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Ebb_{f_0}[\wh{T}] = t_a$, and apply Chebyshev's inequality to deduce
\begin{equation*}
\Pbb_{f_0}\bigl(\wh{T} < t_a\bigr) \leq \Pbb_{f_0}\Bigl(|\wh{T} - \Ebb_{f_0}[\wh{T}]|^2 > |\Ebb_{f_0}[\wh{T}] - t_a|^2\Bigr) \leq \frac{\Var\bigl[\wh{T}\bigr]}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} = \frac{2K/n^2 + 4/n\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2}.
\end{equation*}
Thus we have upper bounded the Type II error by the sum of two terms, each of which are no more than $1/(2b)$, as we now show. For the first term, after noting that~\eqref{pf:fixed_graph_testing_1} implies $\Ebb_{f_0}[\wh{T}] - t_a \geq \sqrt{2K}/n \cdot \sqrt{2/b}$, the upper bound follows:
\begin{equation*}
\frac{2K/n^2}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} \leq \frac{b}{2}.
\end{equation*}
On the other hand, for the second term we use~\eqref{pf:fixed_graph_testing_1} in two ways: first to conclude that $\Ebb_{f_0}[\wh{T}] - t_a \geq 1/2 \cdot \sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2$, and second to obtain
\begin{equation*}
\frac{4\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{n\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} \leq \frac{4\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{n\bigl(\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2/2\bigr)^2} \leq \frac{16}{n\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2} \leq \frac{b}{2}.
\end{equation*}

\section{Analysis of Spectral Series Estimator}
\label{sec:analysis_spectral_series_estimator}
In this section we prove Proposition~\ref{prop:spectral_series_estimation}. As mentioned in our main text, this proof relies on three facts: (i) a continuous embedding of $H_0^s(\mc{X})$ into $\mc{H}^s(\mc{X})$, (ii) the Weyl's Law asymptotic scaling $\lambda_k(\Delta_P) \asymp k^{2/d}$, and (iii) a local Weyl's Law scaling $\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2 \lesssim K$. (Recall that $(\lambda_k(\Delta_P),\psi_k)$ are defined to be solutions to~\eqref{eqn:laplacian_eigenfunction_neumann}). We record the precise results we need in the following three lemmas.

The first result follows immediately from Lemma~17 of \textcolor{blue}{(Dunlop)}. Recall the spectral Sobolev class $\mc{H}^s(\mc{X})$ defined in~\eqref{eqn:spectral_sobolev_space}. Equip $\mc{H}^s(\mc{X})$ with the norm
\begin{equation*}
\|f\|_{\mc{H}^s(\mc{X})}^2 := \sum_{k = 1}^{\infty} \bigl[\dotp{f}{\psi_k}_P\bigr]^2 \cdot [\lambda_k(\Delta_P)]^2.
\end{equation*}
\begin{lemma}[cf. Lemma~17 of \textcolor{blue}{(Dunlop)}]
	\label{lem:zero_trace_spectral_sobolev}
	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$. Then for any $s \in \mathbb{N}$, we have $H_0^s(\mc{X}) \subseteq \mc{H}^s(\mc{X})$. Moreover, there exists a constant $C$ such that for any $f \in H_0^s(\mc{X})$ we have
	\begin{equation}
	\label{eqn:zero_trace_spectral_sobolev}
	\|f\|_{\mc{H}^s(\mc{X})} \leq C\|f\|_{H^s(\mc{X})}.
	\end{equation}
\end{lemma}

The particular version of Weyl's Law we use is also due to \textcolor{blue}{(Dunlop)}. 
\begin{lemma}[cf. Lemma 28 of \textcolor{blue}{(Dunlop)}]
	\label{lem:weyl}
	Suppose Model~\ref{def:model_flat_euclidean}. Then there exist positive constants $c$ and $C$ such that for all $k \geq 2$, 
	\begin{equation}
	\label{eqn:weyl}
	ck^{2/d} \leq \lambda_k(\Delta_P) \leq Ck^{2/d}.
	\end{equation}
\end{lemma}
As we discuss in our main text, the asymptotic scaling $\lambda_k(\Delta_P) \asymp k^{2/d}$ also plays a key role in our analysis of Laplacian eigenmaps. We point out that although Lemma 28 of \textcolor{blue}{(Dunlop)} assumes $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$, the proof uses only the hypotheses of Model~\ref{def:model_flat_euclidean}: that $p$ is bounded away from $0$ and $\infty$, and that $\partial \mc{X}$ is Lipschitz.

Finally, we need a local version of Weyl's Law, due to \textcolor{blue}{(H\"{o}rmander)}.
\begin{lemma}[cf. Theorem 17.5.3 of \textcolor{blue}{H\"{o}rmander}]
	\label{lem:local_weyl}
	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$. Then there exists a positive constant $C$ such that for all $K \in \mathbb{N} \setminus {0}$,
	\begin{equation}
	\label{eqn:local_weyl}
	\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK.
	\end{equation}
\end{lemma}
Translated into our notation, Theorem 17.5.3 of \textcolor{blue}{(H\"{o}rmander)} says that
\begin{equation*}
\sup_{x \in \mc{X}}\biggl\{\sum_{k: \lambda_k(\Delta_P) \leq \lambda_K(\Delta_P)}  \bigl(\psi_k(x)\bigr)^2\biggr\} \leq C \bigl[\lambda_K(\Delta_P)\bigr]^{d/2},
\end{equation*}
and from here~\eqref{eqn:local_weyl} follows from~\eqref{eqn:weyl}. \textcolor{blue}{(H\"{o}rmander)} proves the result only when $\partial \mc{X} \in C^{\infty}$, but the proof goes through unchanged when $\partial \mc{X} \in C^1$, since $\partial \mc{X} \in C^1$ is sufficient to apply the relevant Sobolev embedding theorem (Lemma 17.5.2 of~\textcolor{blue}{(H\"{o}rmander)}) that is the key to proving Lemma~\ref{lem:local_weyl}.

We note that the assumptions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$ used in Lemmas~\ref{lem:zero_trace_spectral_sobolev} and~\ref{lem:local_weyl} can likely be removed. This would allow us to remove the same assumptions in the statement of Proposition~\ref{prop:spectral_series_estimation}. Since this does not change the main point of the proposition, we do not pursue the details further.

\paragraph{Proof (of Proposition~\ref{prop:spectral_series_estimation}).}
We decompose risk into squared bias and variance,
\begin{equation}
\label{pf:spectral_series_estimation_0}
\Ebb \|\wt{f} - f_0\|_P^2 = \Ebb\| \Ebb[\wt{f}]  - f_0\|_P^2 + \Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2.
\end{equation}
Then some standard arguments (which we give below) give the following upper bound:
\begin{equation}
\label{pf:spectral_series_estimation_1}
\Ebb \|\wt{f} - f_0\|_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s} + \frac{K}{n} + \frac{1}{n}\Ebb\Bigl[(f_0(X))^2 \cdot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr].
\end{equation}
The claim of the proposition follows from~\eqref{pf:spectral_series_estimation_1} and Lemmas~\ref{lem:zero_trace_spectral_sobolev}-\ref{lem:local_weyl}, upon taking $K = \floor{(M^2n)^{d/(2s + d)}}$. (Here we use the convention $\sum_{k = 1}^{0}c_k = 0$ for any sequence $(c_k)$.) \qed

The upper bound~\eqref{pf:spectral_series_estimation_1} is to be compared with~\eqref{eqn:estimation_biasvariance}, which gives upper bounds on the (design-dependent) squared bias and variance of Laplacian eigenmaps. We see~\eqref{eqn:estimation_biasvariance} has terms analogous to the first two terms on the right hand side of~\eqref{pf:spectral_series_estimation_1}, but not the third. The relevance of this third term to $L^2$ estimation problems is discussed by~\cite{birge2008}.

\paragraph{Proof (of \eqref{pf:spectral_series_estimation_1}).}
Since $\{\psi_k\}$ are an orthonormal basis of $L^2(P)$, and $f_0 \in H_0^{s}(\mc{X}) \subseteq L^2(P)$, we have $f_0 = \sum_{k = 1}^{\infty} \dotp{f_0}{\psi_k}_P \psi_k$ in $L^2(P)$. This allows us to expand the squared bias in terms of squared Fourier coefficients of $f_0$, and gives the following upper bound,
\begin{equation*}
\|f_0 - \Ebb{\wt{f}}\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}^2 \leq  \frac{1}{\lambda_{K + 1}(\Delta_P)^s} \sum_{k = K + 1}^{\infty} \lambda_{k + 1}(\Delta_P)^s \dotp{f_0}{\psi_k}^2 \leq \frac{\|f_0\|_{\mc{H}^s}}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s}.
\end{equation*}
On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient,
\begin{equation}
\label{pf:spectral_series_estimation_2}
\Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2 = \sum_{k = 1}^{K} \Var\Bigl[\dotp{{\bf Y}}{\psi_k}_n\Bigr] 
\end{equation}
Then from the law of total variance,
\begin{equation*}
 \Var\bigl[\dotp{{\bf Y}}{\psi_k}_n\bigr] = \Var\Bigl[\Ebb[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] + \Ebb\Bigl[\Var[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] = \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr].
\end{equation*}
Finally, $\Ebb \|\psi_k\|_n^2 = \|\psi_k\|_P^2 = 1$, 
\begin{equation*}
\Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] = \frac{1}{n} \Var\Bigl[f_0(X)\psi_k(X)\Bigr] \leq \frac{1}{n} \Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr],
\end{equation*}
and plugging back into~\eqref{pf:spectral_series_estimation_2} yields the claim.

@@

In Lemma~\ref{lem:graph_seminorm_bias}, we decompose the graph Sobolev semi-norm (a V-statistic) into an unbiased estimate of the non-local Sobolev semi-norm (a U-statistic), and a pure bias term. We establish that the pure bias term will be small (in expectation) relative to the U-statistic whenever $\varepsilon$ is sufficiently small.
\begin{lemma}
	\label{lem:graph_seminorm_bias}
	For any $f \in L^2(\mc{X})$, the graph Sobolev semi-norm satisfies
	\begin{equation}
	\label{eqn:graph_seminorm_bias_1}
	\dotp{L_{n,\varepsilon}^sf}{f}_{n} = U_{n,\varepsilon}^{(s)}(f) + B_{n,\varepsilon}^{(s)}(f),
	\end{equation}
	such that $\mathbb{E}[U_{n,\varepsilon}^{(s)}(f)] = (n - s - 1)!/n! \cdot \dotp{L_{P,\varepsilon}^sf}{f}_P$. If additionally $f \in H^1(\mc{X};M)$ and $\varepsilon \geq n^{-1/d}$, then the bias term $B_{n,\varepsilon}^{(s)}(f)$ satisfies
	\begin{equation}
	\label{eqn:graph_seminorm_bias_2}
	\mathbb{E}\bigl[|B_{n,\varepsilon}^{(s)}(f)|\bigr] \leq \frac{C\varepsilon^2}{\delta n\varepsilon^{2 + d}}M^2.
	\end{equation}
\end{lemma}
Then~\ref{pf:graph_seminorm_ho_1} follows immediately from Lemma~\ref{lem:graph_seminorm_bias}, by Markov's inequality.
\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias}).}
We begin by introducing some notation. We will use bold notation $\bj = (j_1,\ldots,j_s)$ for a vector of indices where $j_i \in [n]$ for each $i$. We write $[n]^s$ for the collection of all such vectors, and $(n)^s$ for the subset of such vectors with no repeated indices. Finally, we write $D_if$ for a kernelized difference operator,
\begin{equation*}
D_if(x) := \bigl(f(x) - f(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
\end{equation*}
and we let $D_{\bj}f(x) := \bigl(D_{j_1}\circ \cdots \circ D_{j_s}f\bigr)(x)$.

@@

It remains to show~\eqref{eqn:graph_seminorm_bias_2}. By adding and subtracting $f(X_{\bj_1})$, we obtain by symmetry that
\begin{equation*}
\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i) = \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot \bigl(f(X_i) - f(X_{\bj_1})\bigr),
\end{equation*}
and consequently
\begin{equation*}
\Ebb\Bigl[\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i)\Bigr] \leq \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr].
\end{equation*}
In Lemma~\ref{lem:graph_seminorm_bias2}, we show that if $f \in H^1(\mc{X};M)$, then for any $i\bj \in [n]^{s + 1}$ which contains a total of $k + 1$ distinct indices, 
\begin{equation*}
\Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr] \leq C_1 \varepsilon^{2 + kd} M^2.
\end{equation*}
This shows us that the expectation of $|B_{n,\varepsilon}^s(f)|$ can bounded from above by the sum over several different terms, as follows:
\begin{align*}
\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] & \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2 \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \frac{1}{(n\varepsilon^d)^s}  \varepsilon^{(|i\bj| - 1)d} \\
& \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2  \sum_{k = 1}^{s - 1} \frac{(n\varepsilon^d)^k}{(n\varepsilon^d)^s}n.
\end{align*}
Finally, we note that by assumption $n\varepsilon^d \geq 1$, so that in the above sum the factor of $(n\varepsilon^d)^k$ is largest when $k = s- 1$. We conclude that
\begin{equation*}
\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] \leq C_1 (s - 1) \frac{\varepsilon^2}{n\varepsilon^{2s + d}}M^2,
\end{equation*}
which is the desired result.

@@

Now, we use the results of the preceding two sections to prove~\eqref{pf:graph_seminorm_ho_2}. We will divide our analysis in two cases, depending on whether $s$ is odd or even, but before we do this we state some facts that will be applicable to both cases. First, we  recall that $L_{P,\varepsilon}$ is self-adjoint in $L^2(P)$, meaning $\dotp{L_{P,\varepsilon}f}{g}_{P} = \dotp{f}{L_{P,\varepsilon}g}_{P}$ for all $f, g \in L^2(P)$. We also recall the definition of the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X})$,
\label{eqn:dirichlet_energy}

@@

\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}^{k + 1}f}{L_{P,\varepsilon}^{k}f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}).

@@

\textcolor{red}{(TODO)}

\subsection{Kernel smoothing bias}
\label{subsec:pf_kernel_smoothing_bias}
